MLH1 Data Report

### 2019-12-18 doc/Main_Report_Results.rmd
updating chapter 2 results

```{r rand, echo=FALSE, include=FALSE, eval=FALSE}

str(AP_mouse_table_w.Ages)

@@

@@

todo: re-run the analysis with MOLF data
make model with strain as fixed effect (since, there are interesting strain effects)

The mixed model lets us separate effects to see how they effect a mouse mean CO count. This separation can also serve as proxies for divergence and polymorphism.

```{r mixed.models.setup, echo=FALSE}
#make are the DFs for the models
# seperate the data, both sexes, only Mus
# decide on Musc - Molossinus
AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x

female.musc.mice.table$mouse <- "1jan20_MOLF_f99"
female.musc.mice.table$DOB <- as.Date("1/1/2020")

female.musc.mice.table$strain <- "MOLF"
female.musc.mice.table$subsp <- "Musc"
female.musc.mice.table$species <- "M.musculus"

female.musc.mice.table$female.stage <- "emb"
female.musc.mice.table$age.days <-""
female.musc.mice.table$age.weeks <- ""
female.musc.mice.table$ID <- ""
female.musc.mice.table$notes <- ""
female.musc.mice.table$euth.date <- ""

DF.HetC.MixedModel <- rbind(DF.HetC.MixedModel, female.musc.mice.table)

DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ.mol[!(is.na(DF.HetC.MixedModel.HQ.mol$mouse) | DF.HetC.MixedModel.HQ.mol$mouse==""), ]
```

what are the table above this?

For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species. (maybe I'll look at Species level with spicilegus...?)

subspecies levels: 2, Musc and Domesticus
strain levels: Dom: 3, Musc: 4
sex: 2 female male

List of mixed models with mean_co

1. mean_co (chr coded)
2. mean_co.HQ
3. mean_co with Mol
4. sub-sampling

```{r HQ.MixedModel, echo=FALSE}
table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)

#mean_co.HQ  (this might remove some of the MOLF males)
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel[DF.HetC.MixedModel$Ncells > 9,]
#DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[DF.HetC.MixedModel.HQ$age.weeks < 25,]
DF.HetC.MixedModel.HQ <- DF.HetC.MixedModel.HQ[!(is.na(DF.HetC.MixedModel.HQ$mouse) | DF.HetC.MixedModel.HQ$mouse==""), ]

#these remove ALL male MSM!
#table(DF.HetC.MixedModel.HQ$strain, DF.HetC.MixedModel.HQ$sex)

summary(HetC.MixedModel.HQ)
ranef(HetC.MixedModel.HQ)

#reduced models
Reduced.strain.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
summary(Reduced.strain.HQ)
ranef(Reduced.strain.HQ)

Reduced.all.HQ <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ)
summary(Reduced.all.HQ)
#ranef(Reduced.all.HQ)

#strain = fixed
strain.fixed.full <- lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
summary(strain.fixed.full)
#when strain added as fixed, MSM, PWD and WSB are signifcant, subsp*sex is most significant
#sexmale:strainWSB p=  0.006740 **
#sexmale  p =  3.98e-05 ***
#strainWSB   p=0.000441 ***
#subspMusc:sexmale  p=  0.069493 . 
#non of the 3-way interactions are significant, very puzzeled that WSB is significant and non of ht ehigh Musc are

strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
summary(strain.fixed)

#sexmale  p= 3.98e-05 ***
#strainKAZ p=  0.052420 .  
#strainWSB  p= 0.000441 ***
#sexmale:strainMSM  p=3.42e-09 ***
#sexmale:strainPWD  p= 5.79e-08 ***
#sexmale:strainSKIVE p= 0.069493 .  
#sexmale:strainWSB  p=  0.006740 **

#maybe this WSB effec is just in the HQ data set?

#anova for comparing reduced and full models -- 1 of the models should be strain as fixed instead
full2strain <- anova(HetC.MixedModel.HQ, Reduced.strain.HQ)#slightly significant

Reading these outputs is not the clearest -- try just printing out the coefficient table with p values.

Mostly matches the base model results. the same fixed effects are significant in the reduced models. the random effects make sense

random effects were tested with 
 #rand LRT
this function might not support MM, (and needs linear models instead)

also throws an error about only looking at 1 random effect

(what does the significant random test mean) -- variance is not 0?

## Mixed Model for the variance

>>>>

```{r rand, echo=FALSE, include=FALSE, eval=FALSE}
#str(AP_mouse_table_w.Ages)

@@

```{r table.for.plot, echo=FALSE}

AP_strain.table.HQ <- ddply(MLH1_data, c("strain", "sex"), summarise,
                        Nmice =length(unique(mouse)),
                        Ncells  =length(nMLH1.foci),
                        mean_co =as.numeric(format(round(mean(nMLH1.foci), 3 ), nsmall=3) ),
                        cV = cv(nMLH1.foci, na.rm = TRUE),
                        var = as.numeric(format(round(   var(nMLH1.foci, na.rm = TRUE),3), nsmall=3)), 
                        sd   = round(sd(nMLH1.foci, na.rm = TRUE), 3),
                       #Standard error (mice vs cells)
                        se  = round(sd / sqrt(Ncells), 3)
    )

```{r Fig1, echo=FALSE, warning=FALSE}

Fig1_v1_strain <- ggplot(AP_strain.table.HQ, aes(y=mean_co, x=strain, color=strain, shape=sex))+ 
  ylab("Average MLH1 Foci per Cell")+

  geom_point(aes(shape = sex), size=3, position=position_dodge(.5))+
  geom_errorbar(aes(ymin=mean_co-(se*2), ymax=mean_co+(se*2) ), width=.3, position = position_dodge(.5), size=1.001)+

  ylim(c(18, 35))+
 scale_color_manual(values=colors_of_strains)+ 
  theme(legend.position="none",
  panel.grid.major = element_blank(), panel.grid.minor = element_blank(),
  panel.background = element_blank(), axis.line=element_line(colour = "black"),
  axis.title.x=element_blank(),axis.text.x=element_blank(),axis.ticks.x=element_blank())+
#remove y labfor male 
  #segments
  annotate("segment", x=.75, xend=4.5, y=19.5, yend=19.5, colour="black", size=1)+
  #Musc
  annotate("segment", x= 5.5, xend=10.2, y=19.5, yend=19.5, colour="black", size=1)+
  #mol
  annotate("segment", x=11, xend=12.25, y=19.5, yend=19.5, colour="black", size=1)+
  #cast
  annotate("segment", x=13, xend=14, y=19.5, yend=19.5, colour="black", size=1)+
  #outgroups
  annotate("segment", x=14.5, xend=16, y=19.5, yend=19.5, colour="black", size=1)+
  #label
  annotate("text", x = c(3,8.5,10,12, 14), #list of all x points 
           y = c(18.5,18.5,33,18.5, 18.5),

   label = c("M. m. domesticus","M. m. musculus", "M. m. molossinus", "M. m. castaneus", "Outgroups"),
           fontface = 'italic', hjust = 0.5, size=3.5)

Fig1_v1_strain
```

These are the basic Mixed Models we used as a frame work to analyze all the MLH1 data.

$$mouse .mean CO ~=~ subsp * sex + rand(strain* sex)$$

$$mouse .var CO ~=~ subsp * sex + rand(strain* sex)$$

The mixed model lets us separate effects to see how they effect a mouse mean CO count. The subspecies effect is a proxy for **Divergence** and the random strain effect is a proxy for **Polymorphism**.

```{r mixed.models.setup, echo=FALSE, warning=FALSE}

#why merged 
AP_mouse_table_w.Ages$strain <- AP_mouse_table_w.Ages$strain.x
AP_mouse_table_w.Ages$sex <- AP_mouse_table_w.Ages$sex.x
AP_mouse_table_w.Ages$subsp <- AP_mouse_table_w.Ages$subsp.x

#female.musc.mice.table$mouse <- "1jan20_MOLF_f99"
#female.musc.mice.table$DOB <- as.Date("1/1/2020")

#female.musc.mice.table$strain <- "MOLF"
#female.musc.mice.table$subsp <- "Musc"
#female.musc.mice.table$species <- "M.musculus"

#female.musc.mice.table$female.stage <- "emb"
#female.musc.mice.table$age.days <-""
#female.musc.mice.table$age.weeks <- ""
#female.musc.mice.table$ID <- ""
#female.musc.mice.table$notes <- ""
#female.musc.mice.table$euth.date <- ""

#DF.HetC.MixedModel <- rbind(DF.HetC.MixedModel, female.musc.mice.table)

DF.HetC.MixedModel.HQ.mol <- DF.HetC.MixedModel.HQ.mol[!(is.na(DF.HetC.MixedModel.HQ.mol$mouse) | DF.HetC.MixedModel.HQ.mol$mouse==""), ]

For mouse averages of MLH1 counts, only strain with observations for both sexes and are from Mus species. (maybe I'll look at Species level with spicilegus.?)

subspecies levels: 2, Musc and Domesticus

strain levels: Dom: 3, Musc: 4

sex: 2 female male

Above are the total number of mice in the data set

These are the List of mixed models I built

1. mean_co (chr coded, instead of factors)

2. mean_co.HQ

3. mean_co with Mol (Mol coded as a distinct subsp)

4. sub-sampling (not sure if I made this one)

```{r HQ.MixedModel.1, echo=FALSE}
#table(DF.HetC.MixedModel$strain, DF.HetC.MixedModel$sex)

#summary(HetC.MixedModel.HQ)
#ranef(HetC.MixedModel.HQ)
```

```{r MixedModel.2, echo=FALSE}
#reduced models
Reduced.strain.HQ <- lme(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ, random = ~1|strain)
#summary(Reduced.strain.HQ)
#ranef(Reduced.strain.HQ)
```

```{r MixedModel.2.2, echo=FALSE}

Reduced.all.HQ <- gls(mean_co ~ subsp * sex, data=DF.HetC.MixedModel.HQ)
#summary(Reduced.all.HQ)
#ranef(Reduced.all.HQ)
```

```{r MixedModel.3, echo=FALSE}
#strain = fixed
strain.fixed.full <- lm(mean_co ~ subsp * sex * strain, data=DF.HetC.MixedModel.HQ)
#summary(strain.fixed.full)
#when strain added as fixed, MSM, PWD and WSB are signifcant, subsp*sex is most significant

#non of the 3-way interactions are significant, very puzzeled that WSB is significant and non of ht ehigh Musc are 
```

```{r MixedModel.4, echo=FALSE}

strain.fixed <- lm(mean_co ~ sex * strain, data=DF.HetC.MixedModel.HQ)
#summary(strain.fixed)

#maybe this WSB effec is just in the HQ data set?
#anova for comparing reduced and full models -- 1 of the models should be strain as fixed instead
full2strain <- anova(HetC.MixedModel.HQ, Reduced.strain.HQ)#slightly significant

```{r try.blah, echo=FALSE}
#library(jtools)
#states <- as.data.frame(state.x77)
#fit <- lm(Income ~ Frost + Illiteracy + Murder, data = states)
#summ(fit)
#this is a nice display package for lm

res.table <- as.data.frame(coef(summary(Reduced.all.HQ)))
res.table
```

## Mixed Model for variance

@@
